IRTF Spectro-Photometry of Two Rotating Brown Dwarfs: \\2MASS J08354256-0819237 and 2MASS J18212815+1414010

### 2016-11-20 ms.tex
I went through and updated the section on the amplitude spectrum fits.

We fit the time series of each brown dwarf in 25 spectral bins each to assess the wavelength-dependent nature of their variations.
The simple model can be described as 
\begin{equation}
F(t) = A * \sin(2 \pi (t - t_0)/\tau) + B t + C
\end{equation}
where $F$ is the flux as a function of time ($t$), A is the amplitude of the variations, $t_0$ is the time offset of the variations and $\tau$ is the and B and C are linear terms describing the baseline trend.
For the fits, we include a 4 free parameters: $A$, $t_0$, $B$ and $C$.
We fixed the sinusoidal period $\tau$ at the literature value of periodicity because there are also variations due to variable telluric variation that are not relevant to the brown dwarf system.
We imposed a prior that $A \geq 0$ so that a solution with a $\pi$ phase offset and negative amplitude isn't found.

\begin{figure*}[!]
\centering
\subfloat[2MASS J0835 - \colorbox{red}{fit might need more work} ]{
	\includegraphics[width=0.5\textwidth]{amp_vs_wavl_j0835.pdf}
	\label{fig:ampspec0835}
	}
\subfloat[2MASS J1821]{
	\includegraphics[width=0.5\textwidth]{amp_vs_wavl_j1821.pdf}
	\label{fig:ampspec1821}
	}
	\caption{Sine fit amplitude versus wavelength for each brown dwarf. The regions at 1.4~$\mu$m and 1.9~$\mu$m (shaded with a salmon color) are contaminated by telluric absorption so the variations can likely be due to telluric variability not corrected by the reference star.}
	\label{fig:ampSpec}
\end{figure*}

Figure \ref{fig:ampSpec} shows the best fit amplitudes as a function of wavelength.
We use the standard deviation of points near minimum to estimate the noise in the flux time series.
The Levenberg-Marquardt chi-squared minimization code \texttt{mpfit} is used to fit each wavelength bin independently and propagate the errors in the flux time series to amplitude uncertainty.
The amplitude spectrum of 2MASS J0835 shown in Figure \ref{fig:ampspec0835} shows random fluctuations as a function of wavelength.
2MASS J1821, on the other hand, shows a clear decreasing trend as function of wavelength.
This was also observed in the shorter wavelength range of 1.1~$\mu$m to 1.7~$\mu$m in \citet{yang2016exStormsBD}.

This type of wavelength trend was also recently observed for WISEP J004701.06+680352.1 \citep{lew2016w0047}, which has a much $J-K$ 2.55 spectrum.
In WISEP J0047, the wavelength trend could be fit by $\sim 0.4~\mu$m forsterite grains.

@@

We model the wavelength dependence of 2MASS 1821 as an optically thin cloud layer consisting of a log-normal distribution of particle sizes.
In this simple model, the amplitude of the variability is directly proportional to the opacity or $Q_{ext}$ of the particles.
We model the extinction coefficient, $Q_{ext}$ from the \texttt{IDL} Mie theory code \texttt{mie\_single.pro} \citep{grainger04} to calculate the extinction as a function of wavelength.
We use the complex indices of refraction from \citet{dorschner95pyrox} for Mg-rich pyroxene grains (MgSiO$_3$).
For the particle size distribution we assume a log-normal distribution around a characteristic size with a distribution $\sigma$ of 0.5.

@@

The WFC3 amplitude spectrum is shown with the data in Figure \ref{fig:ampspec1821mie}.
There is a multiplicative offset between the IRTF and WFC3 spectrum because they were taken at different epochs and therefore different spatial cloud and temperature distributions.
The WFC3 spectrum is not affected by the telluric absorption the way the SpeX data is.
\citet{yang2015hstRotBD} find that the WFC3 spectrum has no significant water vapor absorption which indicates a high haze layer in 2MASS J1821, consistent with our Mie scattering fit.

@@

Another simple model we attempt is a spot model - that the brown dwarf spectral variability has to do with a non-uniform temperature distribution on the brown dwarf in the same way that starspots modulate the brightness of hydrogen burning stars as they spin.
In Figure \ref{fig:ampspec1821tdiff}, we fit the same amplitude spectrum as in Figure \ref{fig:ampspec1821mie}.
This time we model the data with two temperature regions as shown in Figure \ref{fig:tdiffschem} with two opposite longitudes of the planet at the brightness minimum and maximum.
We assume that are each in local thermodynamic equilibrium (LTE) so that the flux is proportional to the Planck function times the areal coverage.
The flux ratio between maximum and minimum flux is then
\begin{equation}
f_{R,\lambda} = \frac{\alpha_1 B_\lambda(T_1) + (1-\alpha_1) B_\lambda(T_2)}{\alpha_2 B_\lambda(T_1) + (1-\alpha_2) B_\lambda(T_2)}
\end{equation}
where $\alpha_1$ and $\alpha_2$ are the areal fractions of the two temperature layers limited to [0,1], $B_\lambda$ is the Planck function, and $T_1$ and $T_2$ are the temperatures of the two levels.
Using equation \ref{eq:ampFromRatio}, the amplitude then becomes
\begin{equation}
A_\lambda = \frac{\left(\alpha_1 - \alpha_2 \right) \left(B_\lambda(T_1) - B_\lambda(T_2) \right)}{\left(\alpha_1 + \alpha_2\right) B_\lambda(T_1) + \left(2 - \alpha_1 - \alpha_2\right) B_\lambda(T_2)}.
\end{equation}
If the temperature difference is significant so that one Planck function is in the Wien's limit while the other is in the Rayleigh Jeans tail, the cooler temperature essentially becomes black and the contrast approaches a constant for short wavelengths.
If we ignore projection effects for small area coverage, the effective temperature is 
\begin{equation}
\sigma T_{eff}^4 \approx (\alpha_1 + \alpha_2) \sigma T_1^4 + (2 - \alpha_1 - \alpha_2) \sigma T_2^4
\end{equation}
where $\sigma$ is the Stephan-Boltzmann constant.

Our model best-fits the data with $\alpha_i << $1 and also with much lower temperatures than equilibrium, shown in Figure \ref{fig:ampspec1821tdiff}.
This model has a reduced chi-squared of 1.3, consistent with the data, but not with the effective temperature.
If we restrict the model to be a two-faced model where $\alpha_1$ = 1 and $\alpha_2$ = 0, so that the two halves of the brown dwarf are different temperatures, the best-fit is a poor match to the data, also shown in Figure \ref{fig:ampspec1821tdiff} with a $\bar{\chi}^2$ = 15.
Finally, we attempt a model where we restrict the temperature limits to [1300,2500] to ensure that the final effective temperature is close to the measured 1600 K \citep{gagne2015banyan7}.
The best fit model has $T_{eff} \approx $1590 K, but pushes to the edges of our prior with $T_1$ = 2500 K and $T_2$ = 1300 K.
The temperature contrasts in these models (with the exception of the two-faced model which poorly matches the data) are significantly more than the expected 50 K horizontal temperature variations expected in circulation models \citep{showman2013bdgpDynamics}.
Therefore, we suspect that hot or cold spots on the surface a are a less likely scenario to explain the variability of 2MASS 1821.

@@

We observed two brown dwarfs of spectral type L4-L5: 2MASS 1821 and 2MASS 0835.
These brown dwarfs have temperatures near the condensation temperatures of MgSiO$_3$ and Mg$_2$SiO$_4$.
We obtained high precision ground-based spectrophotometry of both systems to measure the variability spectra of these systems with the IRTF.
After fitting sinusoids to both of the time series, we find that 2MASS 0835 has low levels of variability and not wavelength dependence.

>>>>

We fit the time series of each brown dwarf in 25 spectral bins each to assess the wavelength-dependent nature of their variations.
The simple model can be described as 
\begin{equation}
F(t) = A * \sin(2 \pi (t - t_0)/\tau) + B t + C
\end{equation}
where $F$ is the flux as a function of time ($t$), A is the amplitude of the variations, $t_0$ is the time offset of the variations and $\tau$ is the and B and C are linear terms describing the baseline trend.
For the fits, we include a 4 free parameters: $A$, $t_0$, $B$ and $C$.
We fixed the sinusoidal period $\tau$ at the literature value of rotation period (Table \ref{tab:bdProp}) because are trends (with a period of $\sim$24 hr) due to uncorrected telluric variations that that are not relevant to the brown dwarf system.
We imposed a prior that $A \geq 0$ so that a solution with a $\pi$ phase offset and negative amplitude isn't found.

\begin{figure*}[!]
\centering
\subfloat[2MASS J0835 - \colorbox{red}{fit might need more work} ]{
	\includegraphics[width=0.5\textwidth]{amp_vs_wavl_j0835.pdf}
	\label{fig:ampspec0835}
	}
\subfloat[2MASS J1821]{
	\includegraphics[width=0.5\textwidth]{amp_vs_wavl_j1821.pdf}
	\label{fig:ampspec1821}
	}
	\caption{Sine fit amplitude versus wavelength for each brown dwarf. The regions at 1.4~$\mu$m and 1.9~$\mu$m (shaded with a salmon color) are contaminated by telluric absorption so the variations can likely be due to telluric variability not corrected by the reference star. 2MASS J0835 shows no clear trend in wavelength, while 2MASS J1821 has a strong slope indicative of Mie-scattering particles.}
	\label{fig:ampSpec}
\end{figure*}

Figure \ref{fig:ampSpec} shows the best fit amplitudes as a function of wavelength.
We use the standard deviation of points near minimum to estimate the noise in the flux time series.
The Levenberg-Marquardt chi-squared minimization code \texttt{mpfit} is used to fit each wavelength bin independently and propagate the errors in the flux time series to amplitude uncertainty.
The amplitude spectrum of 2MASS J0835 shown in Figure \ref{fig:ampspec0835} shows small ($\lesssim 0.4\%$) amplitudes with a flat trend as a function of wavelength.
2MASS J1821, on the other hand, shows a clear decreasing trend as function of wavelength.
This was is the same trend observed for 2MASS J1821 by HST over a shorter wavelength range (1.1 to 1.7~$\mu$m) by \citet{yang2016exStormsBD}.

@@

We model the wavelength dependence of 2MASS 1821 as an optically thin cloud layer consisting of a log-normal distribution of particle sizes.
In this simple model, the amplitude of the variability is directly proportional to the opacity or $Q_{ext}$ of the particles.
We model the extinction coefficient, $Q_{ext}$ from the \texttt{IDL} Mie theory code \texttt{mie\_single.pro} \citep{grainger04} to calculate the extinction as a function of wavelength.
We use the complex indices of refraction from \citet{dorschner95pyrox} for Mg-rich pyroxene grains (MgSiO$_3$).
For the particle size distribution we assume a log-normal distribution around a characteristic size, $s$ and a standard deviation of the logarithm (scale parameter), $\sigma$ =0.5.

@@

The WFC3 amplitude spectrum is shown with the data in Figure \ref{fig:ampspec1821mie}.
There is a multiplicative offset between the IRTF and WFC3 spectrum because they were taken at different epochs and therefore different spatial cloud and temperature distributions.
The WFC3 spectrum is not affected by the telluric absorption the way the SpeX data is.
\citet{yang2015hstRotBD} find that the WFC3 spectrum has no significant water vapor absorption which indicates a high haze layer in 2MASS J1821 and consistent with our Mie scattering fit.

A similar amplitude versus wavelength trend was recently observed for WISEP J004701.06+680352.1 \citep{lew2016w0047}, which has very red ($J-K$ = 2.55) spectral energy distribution thought to be caused by a high altitude haze layer.
In WISEP J0047, the wavelength trend could be fit by $\sim 0.4~\mu$m forsterite grains, similar to our characteristic particle size.

@@

Another simple model we attempt is a spot model - that the brown dwarf spectral variability has to do with a non-uniform temperature distribution on the brown dwarf in the same way that starspots modulate the brightness of hydrogen burning stars as they spin.
In Figure \ref{fig:ampspec1821tdiff}, we fit the same amplitude spectrum as in Figure \ref{fig:ampspec1821mie}.
This time we model the data with two temperature regions as shown in Figure \ref{fig:tdiffschem} with two opposite longitudes of the planet at the brightness minimum and maximum.
We assume that are each in local thermodynamic equilibrium (LTE) so that the flux is proportional to the Planck function times the areal coverage.
The flux ratio between maximum and minimum flux is then
\begin{equation}
f_{R,\lambda} = \frac{\alpha_1 B_\lambda(T_1) + (1-\alpha_1) B_\lambda(T_2)}{\alpha_2 B_\lambda(T_1) + (1-\alpha_2) B_\lambda(T_2)}
\end{equation}
where $\alpha_1$ and $\alpha_2$ are the areal fractions of the first ($T_1$ and second ($T_2$) regions respectively, $B_\lambda$ is the Planck function, and $T_1$ and $T_2$ are the temperatures of the two levels.
$\alpha_1$ and $\alpha_2$ are limited to [0,1].
Using equation \ref{eq:ampFromRatio}, the amplitude then becomes
\begin{equation}
A_\lambda = \frac{\left(\alpha_1 - \alpha_2 \right) \left(B_\lambda(T_1) - B_\lambda(T_2) \right)}{\left(\alpha_1 + \alpha_2\right) B_\lambda(T_1) + \left(2 - \alpha_1 - \alpha_2\right) B_\lambda(T_2)}.
\end{equation}
If the temperature difference is significant so that one Planck function is in the Wien's limit while the other is in the Rayleigh Jeans tail, the cooler temperature essentially becomes black and the contrast approaches a constant for short wavelengths.
If we ignore projection effects for small area coverage, the effective temperature is 
\begin{equation}
\sigma T_{eff}^4 \approx (\alpha_1 + \alpha_2) \sigma T_1^4 + (2 - \alpha_1 - \alpha_2) \sigma T_2^4
\end{equation}
where $\sigma$ is the Stephan-Boltzmann constant.

We start by considering a two-faced model where $\alpha_1$ = 1 and $\alpha_2$ = 0, so that the two halves of the brown dwarf are different temperatures.
To be close to the effective temperature of the brown dwarf, we fix $T_1$ to be 1600 K and only vary $T_2$.
This model is a poor match to the data, also shown in Figure \ref{fig:ampspec1821tdiff} with a reduced chi-squared, $\bar{\chi}^2$ = 15.
Next, we attempt a model where we allow $\alpha_1$ and $\alpha_2$ to be free parameters and restrict the temperature limits of the two regions to [1300,2500] to ensure that the final effective temperature is close to the measured 1600 K \citep{gagne2015banyan7}.
The best fit model has $T_{eff} \approx $1590 K and $\bar{\chi}^2$=2.1, but pushes to the edges of our prior with $T_1$ = 2500 K and $T_2$ = 1300 K.
Finally, we consider a model where all four parameters, $\alpha_1$, $\alpha_2$, $T_1$ and $T_2$ are all free.
With these loose constraints, the model fits the data with $\bar{\chi}^2$ = 1.3, consistent with the data, but not with the effective temperature ($T_{eff}$ = 850 K).
THe $\alpha$ values are much smaller indicating that the model is like dark patch.

The temperature contrasts in these models (with the exception of the two-faced model which poorly matches the data) are significantly larger than the expected 50 K horizontal temperature variations expected in circulation models \citep{showman2013bdgpDynamics}.
Therefore, we suspect that hot or cold spots on the surface a are a less likely scenario to explain the variability of 2MASS 1821.

@@

We observed two brown dwarfs of spectral type L4-L5: 2MASS 1821 and 2MASS 0835.
These brown dwarfs have temperatures near the condensation temperatures of MgSiO$_3$ and Mg$_2$SiO$_4$.
We obtained high precision ground-based spectrophotometry of both systems to measure the variability spectra of these systems with the IRTF.
After fitting sinusoids to both of the time series, we find that 2MASS 0835 has low levels of variability and no significant wavelength dependence.
